\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-17 trunk/europa.tex
in europa.tex: fleshed out inference section, finished search section.

\eu is implemented as a library that encapsulates all the elements
described above and makes them easily accessible through modeling and
API layers.  A user will typically use the modeling facilities to
inject a problem domain description, and the API to extract and
manipulate plans, and to obtain information about plan flaws and
constraint violations at any stage of the planning process.

@@

The \textbf{Constraint Reasoning Engine} (CRE) manages domain types,
variables and constraints that define relationships among them. It
also provides an efficient arc consistency mechanism
\cite{mackworth77}. The CRE is designed so that specialized reasoning
algorithms for specific constraints can be easily and efficiently
plugged in.

@@

The \textbf{Rules Engine} (RE) modularly adds the capability to
support domain rules that describe dependencies between tokens,

@@

The \textbf {Language Modules}: Describing problem domains and problem
instances in terms of objects, variables and domain rules in a
programmatic way is cumbersome to the extent of being impractical,
that motivated the creation of modeling languages that support those
definitions in a much more natural and compact way. NDDL is the main
modeling language for \eu and will be described in detail in the
Modeling section below. ANML is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

In an effort to mimimize the amount of work that a user needs to
perform to solve specific problems, \eu also provides extension
modules that have been found to be useful in many domains:

@@

  starts(effect Own purchase);
  eq(purchase.product,product);
}
\end{verbatim}

@@

Now, having that we have a specification for the Shopping Agent
domain, here is how a particular problem instance could be specified:

@@

A problem instance consists mainly of

@@

\item \textbf{Goals}: Goals are represented as tokens in \eu's Plan
  Database, differently from Facts, Goals trigger inference in \eu
  which determines whether flaws (unresolved subgoals, overloaded
  resources, etc., flaws will be covered in detail in the search
  section below) need to be resolved, if that is the case, search
  modules can be invoked which will look for a plan where all the
  flaws are resolved.

@@

\comment{explain how inference is used to detect perform bounds
  propagation, domain reduction and to detect constraint violations.} 
- talk about AC3 implementation and extension mechanisms
- talk about europa's constraint library

The Temporal Network: The notion of time is central to temporal
planning. \eu uses variables to explicitly represent timepoints for
plan activities and states. Constraints among timepoints provide a
natural way to express domain axioms. For example, in order to state
that activity A must occur before activity B we can say that the end
timepoint of A is $leq$ the start timepoint of B.

@@

\cite{dechter91} proposed that constraints among timepoints can be
grouped together to form a Simple Temporal Network (STN). Such a
network can be transformed into a Distance Graph (DG) where the
outward arc from a node represents the maximum distance from the
source node to the target node. They also showed that shortest path
algorithms could be used to propagate values in the network and
discover a negative cycle. A negative cycle is a path from a node to
itself that has a path length less than 0. If such a cycle exists, the
network is inconsistent. It was further shown that a single-source
shortest path algorithm was sufficient to detect a negative cycle and
provide sufficient propagation to yield a backtrack-free search. Thus
we have an efficient and complete algorithm for propagating an
STN. These results build on the already established notion of a CSP
and are naturally incorporated into the general representation and
propagation scheme used in \eu.  The diagram below illustrates a
simple STN with just 2 variables and a single constraint. It also
shows the resulting DG.

@@

% \comment{ add figure. Use Europa wiki figure; or use conor figure from slides}

Resources:

@@

\item \textbf{Reservoir} resources can have both 'consume' and
  'produce' tokens. Note that these tokens do NOT have the usual
  start/end/duration token variables, since occur at a single instant
  in time. Instead, they have a single variable, 'time' used to
  represent that instant.

\item \textbf{Reusable} resources have 'uses' tokens that use a
  quantity for the duration of the token. These tokens do have the
  usual start/end/duration variables.

\item \textbf {Unary} resources are reusable resources with unit
  quantity and have 'use' tokens (again, with the usual
  start/end/duration variables).
\end{enumerate}

\comment{talk about resource envelope computation by using MaxFlow formulation.}

@@

\item A problem instance (also called initial state) that consists of:
\item Variable and Object instances that exist for the entire planning
  horizon.
\item Temporally scoped predicate and action instances.
\item Temporally scoped goals.

@@

\comment{expand a little more on how heuristics can be implemented by
  applying ordering/filtering for chooseFlaw and makeDecisionPoint}
examples:
- deal with earliest flaws first
- deal with planning (open condition flaws) before scheduling (threats) ones

It should be emphasized that while this algorithm is commonly
employed, it is only one of many that could be
implemented. \comment{give examples of local search solvers for
  scheduling and CSP?}

N-Queens solver:
\begin{verbatim}
    public void solve(int maxIter)
    {
    	init();

    	for (int i=0;psengine_.getViolation() > 0 && i < maxIter;i++) {
    	    PSVariable queenToMove = getQueenWithMaxViolation();
    		int curPos = queenToMove.getSingletonValue().asInt();

    	    boolean moved = false;
            SortedSet<Move> moves = getMoves(queenToMove,curPos);             
            for (Move m : moves) {
            	moved = makeMove(queenToMove,curPos,m,false);
            	if (moved)
            	    break;
            }

            if (!moved) 
            	makeMove(queenToMove,curPos,moves.first(),true);

            checkSolution(); // See if we have a new best solution            
            notifyIterationCompleted(curIteration_++);

            if (curIteration_-bestIter_ > 50) 
            	restart();
    	}
    }
\end{verbatim}

RCPSP solver \comment{Do we need this example?}:
\begin{verbatim}
    public void solve(PSEngine psengine,
    		          long timeout, // in msecs
    		          int bound)
    {
       init(psengine,timeout,bound);

       for (int i=0; true ; i++) {
          flatten();
          updateSolution(i);
          updateCriticalPrecedences();

          if ((nbStable_ > maxStable_) || (bestMakespan_ <= makespanBound_))
             break;

          if (timer_.getElapsed() > timeout) {
              timedOut_ = true;
              break;
          }

          relax();
          curIteration_++;
       }

       restoreBestSolution();
    }

\end{verbatim}

\comment{explain how other solvers can be built, maybe use NQueens or
  RCPSP example to illustrate. Or maybe build a little specialized
  solver for shopping or rover examples.}

>>>>

\eu is implemented as a library that encapsulates all the elements
described above and makes them easily accessible through modeling and
Application Programming Interface (API) layers.  A user will typically use the modeling facilities to
inject a problem domain description, and the API to extract and
manipulate plans, and to obtain information about plan flaws and
constraint violations at any stage of the planning process.

@@

The \textbf{Constraint Reasoning Engine} (CRE) manages domain types,
variables and constraints that define relationships among them. It
also provides an efficient arc consistency mechanism
\cite{mackworth77}. The CRE is designed so that specialized reasoning
algorithms for new constraints can be easily and efficiently
plugged in.

@@

The \textbf{Rules Engine} (RE) modularly adds the semantics of domain rules that describe dependencies between tokens.

@@

The \textbf {Language Modules}: Describing problem domains and problem
instances in terms of objects, variables and domain rules in a
programmatic way is cumbersome to the extent of being impractical,
this motivated the creation of modeling languages that support those
definitions in a much more natural and compact way. NDDL is the main
modeling language for \eu and will be described in detail in the
Modeling section below. ANML is a language under development that
provides a more intuitive syntax \cite{smith08}.

@@

In an effort to mimimize the amount of work that a user needs to
perform to solve specific problems, \eu also bundles extension
modules that have been found to be useful in many domains:

@@

  meets(effect Own purchase);
  eq(purchase.product,product);
}
\end{verbatim}

@@

Having that specification for the Shopping Agent
domain, here is how a particular problem instance could be stated:

@@

A problem instance consists of:

@@

\item \textbf{Goals}: Goals are represented as tokens in \eu's Plan
  Database. Differently from Facts, Goals trigger inference in \eu
  which determines whether flaws (unresolved subgoals, overloaded
  resources, etc.) need to be resolved, if that is the case, search
  modules can be invoked which will look for a plan to resolve those
  flaws.

@@

In \eu,  all the Variables and Constraints from a problem domain domain description are kept inside the Constraint Reasoning Engine (CRE).
The CRE provides a framework where any number of constraint types can collaborate to perform inference and prune variable domains using Bounds Propagation and Domain Reduction techniques. \eu uses a Propagator pattern, in which a Propagator object keeps track of a set of constraint instances of the same type and is responsible for ensuring that whenever any variable changes or constraints are introduced or dropped, all related variables are updated through inference.

In a planning problem, variable or constraint changes can come directly from an end user (by introducing new activities or goals and relating them to the rest of the plan through constraints, by changing the start time of an activity, etc), or can come from an automated planner (by expanding subgoals, introducing precedence constraints to resolve mutual exclusion conflicts, etc). When these changes happen, each Propagator is only able to ensure consistency for the constraints that are associated with it.  To maintain global consistency, the CRE implements a variant of the AC-3 algorithm \cite{mackworth77} where it asks Propagators to perform inference until the Constraint Network returns to an stable state.

\comment {add europa's AC-3 algorithm here}

A \eu user can introduce new constraint types by implementing specialized Constraint and Propagator classes in C++, however, \eu provides a built-in with over 25 arithmetic, temporal, set and resource constraint types so that a large number of problem domains can be modeled without having to introduce new constraint types. \comment{insert table from http://code.google.com/p/europa-pso/wiki/ConstraintLibraryReference ? }
Temporal and resource constraints are specially important to model real-life planning domains, so it is worth going into some more detail about how \eu's built-in support works for those constraint types.

\textit{Temporal Constraints:} \eu uses variables to explicitly represent timepoints for
plan activities and states. Constraints among timepoints provide a
natural way to express domain axioms. For example, in order to state
that activity A must occur before activity B we can say that the end
timepoint of A is $leq$ the start timepoint of B.

@@

\cite{dechter91} proposed that constraints among timepoints can be
grouped together to form a Simple Temporal Network (STN). Such a
network can be transformed into a Distance Graph (DG) where the
outward arc from a node represents the maximum distance from the
source node to the target node. 
The diagram below illustrates a simple STN with just 2 variables and a single constraint. It also
shows the resulting DG.

@@

\cite{dechter91} also showed that shortest path
algorithms could be used to propagate values in the network and
discover a negative cycle. A negative cycle is a path from a node to
itself that has a path length less than 0. If such a cycle exists, the
network is inconsistent. It was further shown that a single-source
shortest path algorithm was sufficient to detect a negative cycle and
provide sufficient propagation to yield a backtrack-free search. Thus
we have an efficient and complete algorithm for propagating an
STN. \eu's Propagator for temporal constraints is a straight implementation of this algorithm.

\textit{Resource Constraints:} Most real life planning problems have to deal with 
the fact that plans made up of actions that require resources to be performed typically
have only a limited amount of those resources available (think of fuel, storage capacity, etc).
\eu provides built-in support for resources whose availability can be measured with real numbers, 
in terms of \eu's  representation, resource object types and their related constraints are provided as follows:

@@

\item \textbf{Reservoir} resources can have both \texttt{consume(resource,time,quantity)} and
  \texttt{produce(resource,time,quantity)} constraints. Production and consumption is instantaneous (it happens at a single point in time)

\item \textbf{Reusable} resources have \texttt{uses(resource,start,duration,quantity)} constraints that use a
  quantity of the resource for a specific duration, then release it.

\item \textbf {Unary} resources are Reusable resources with capacity=1, which allows \eu to perform 
some run time savings compared to the generic Reusable class.
\end{enumerate}

\eu offers configuration options so that in addition to the usual constraint violations when a resource is completely depleted, plan flaws can be generated whenever a resource level 
is outside a specific upper or lower limit, or when instantaneous production or consumption exceeds a specific threshold.

Recall that \eu supports a flexible temporal representation, where time is represented by intervals, instead of single values. In this context, determining whether a resource is the source of a plan flaw or a constraint violation in general is more difficult than just computing a profile 
of the resource and checking against some limits. 
In the context of time flexibility, there are many resource profiles that are possible, depending on when exactly each production/consumption event happens out of all the possible values that are in the domain of the variables that represent the time for the event.
\cite{mus02} proposed an algorithm where the variables that represent the time for production/consumption events are placed in a temporal network, similar to the network used by the Temporal Propagator above, then the constraints on the resource can be seen as quantities flowing through that network. Computing the maximum flow will yield the maximum resource usage, and thus a lower bound on the resource level, and computing a minimum flow will similarly yield an upper bound. Since maximum flow on a temporal network can be computed in polynomial time, \eu takes advantage of this algorithm to efficiently compute upper and lower bound envelopes on the resource level.

\comment{provide snapshot of resource envelopes generated by europa}

Resource envelopes can be very useful when trying to figure how brittle a plan in the face of changing activity durations or start times. They can also be useful when searching for plans, although they are relatively expensive to compute so their ability to prune the search space must be weighted against the ability to search faster \comment{reference Barreiro,Morris, Bresina paper on grounded inference}.

@@

\item A problem instance (also called initial state) that consists of:
    \begin{enumerate}
        \item Variable and Object instances that exist for the entire planning
  horizon.
        \item Temporally scoped predicate and action instances.
        \item Temporally scoped goals.
\end{enumerate}

@@

Heuristics can be implemented by applying ordering/filtering policies for
\texttt{chooseFlaw} and \texttt{makeDecisionPoint}, for instance, by dealing with earliest flaws first, 
or by making planning decisions (open condition flaws) before scheduling ones (threats), etc.

It should be emphasized that while this algorithm is commonly employed, it is only one of many that could be
implemented. As a simple example, consider the following solver for the N-Queens problem, where there is an integer variable 
Vi to represent the queen in column i, so that for instance, V1=3 means that the variable on column 1 is placed on row 3. 
non-attack constraints are posted as arithmetic constraints among the Vi variables to ensure that all the rows and diagonals contain only one queen.

\begin{verbatim}
    public void solve(ConstraintEngine ce, int maxIter)
    {
    	curIter=0;

    	while ((ce.getViolations() > 0) && (curIter < maxIter)) {
	    Variable queenToMove = getQueenWithMaxViolation(ce);
    	     int curPos = queenToMove.getCurrentValue();

            SortedSet<int> moves = getMoves(ce,queenToMove,curPos);             
  	   boolean moved = false;
            for (int newPos : moves) {
            	moved = tryMove(queenToMove,curPos,newPos);
            	if (moved)
            	    break;
            }

            if (!moved) 
            	forceMove(queenToMove,curPos,moves.first());

            checkSolution(); // See if we have a new best solution            
            curIteration_++;
    	}
    }
\end{verbatim}

\eu's Constraint Reasoning Engine can be used to compute the non-attack constraint violations (\texttt{ce.getViolations()} is a straight call on the \eu API), 
to easily pick the queen that has he most violations by implementing \texttt{getQueenWithMaxViolation(ce)} and to look ahead and see what 
moves are possible for a queen that would remove its violations by implementing \texttt{getMoves(ce,queenToMove,curPos)}.

In similar fashion, the constraint violation and flaw reporting mechanisms that \eu provides can be used as building blocks that specialized or general purpose search algorithms can use to construct plans.
